---
title: "Chapter 1"
author: "Thomas Sandmann"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Expected frequency of X to be exactly 3 for a Poisson distribution with rate 5:

```{r}
dpois(x = 3, lambda = 5)
```

Probabilities of seeing 0,1,2,...,12 mutations, as modeled by the Poisson(5) distribution. The plot shows that we will often see 4 or 5 mutations but rarely as many as 12. The distribution continues to higher numbers ($13,...$), but the probabilities will be successively smaller, and here we don\'t visualize them.

```{r fig.width = 5.9, fig.height = 5}
barplot(dpois(0:12, 5), names.arg = 0:12, col = "red")
```

```{r fig.cap="Simulated distribution of B(10000, $10^{-4}$) for 300,000 simulations.", fig.width = 7, fig.height = 5.5}
rbinom(1, prob = 5e-4, size = 10000)
simulations = rbinom(n = 300000, prob = 5e-4, size = 10000)
barplot(table(simulations), col = "lavender")
```

🎓 Task: Verify by simulation that the sum of 50 independent Bernoulli variables with `p = 0.01` is – to good enough approximation – the same as 
a `Poisson(0.5)` random variable.

```{r}
set.seed(123)
n_trials = 5e5
sum(rbinom(n = n_trials, size = 50, prob = 0.01))
sum(rpois(n = n_trials, lambda = 0.5))
```

```{r}
set.seed(123)
barplot(rpois(n = 100, lambda = 0.5), names.arg = 1:100, ylim = c(0, 7),
        width = 0.7, xlim = c(-0.5, 100.5))
```

Experiment with the random number generator that generates all possible numbers between 0 and 1 through the function called `runif`. Use it to generate a random variable with 4 levels (A, C, G, T) with probabilities
1/8, 3/8, 3/8 and 1/8, respectively.

```{r}
probabilities <- c(1, 3, 3, 1) / 8
probabilities
bases <- cut(
  runif(1e5),
  breaks = c(0, cumsum(probabilities)),
  labels = c("A", "C", "G", "T"))
table(bases) / 1e5
```
